---
title: "Assignment.demo"
date: "2024-08-27"
output:
  html_document: default
  word_document: default
  pdf_document: default
---

**Loading the data and import libraries**

Introduction to R for Data Visualisation

Install package 'tidyverse'

library dplyr to use function such as group by library ggplot2 to find outline

```{r}
library(tidyverse)
library(dplyr) 
library(ggplot2)
#library(forcats)# For factor manipulation
library(caret)# For model evaluation
library(rpart.plot) # For VIF
library(pROC) #ROC curve
library(ellipse)
library(tree) # classification tree
library(funModeling)
library(patchwork) #combine 3 geaphs
```

Data import from csv.

```{r}
marketing_data <- read.csv("C:/Users/kylew/OneDrive/Desktop/uni/MMA marketing_data_sample.csv")
```

number of rows (observations)

```{r}
nrow(marketing_data)
```

number of columns (variables)

```{r}
ncol(marketing_data)
```

Data type

```{r}
glimpse(marketing_data)
```

### **Data Exploration**

Summary data (only numerical data)

```{r}
numerical_data <- select(marketing_data, where(is.numeric))
summary(numerical_data)
```

if the mean is greater than the median, the distribution is positively skewed.

if the mean is less than the median, the distribution is negatively skewed.

### Handing missing data

Find missing value (no missing value, only unknown values)

```{r}
which(is.na(marketing_data))
sum(is.na(marketing_data))
```

Find any duplicated value - identical elements would respond False

```{r}
table(duplicated(marketing_data))
```

### Data Visualisation outliers

**Age**

Analysis Create a histogram for age by output

```{r}
ggplot(data = marketing_data, mapping = aes(x = age, fill = y)) + labs(title = "Age Distribution by Output") + geom_histogram(binwidth = 5) + theme_minimal() + theme(plot.title = element_text(hjust = 0.5, size = 16, face ="bold"),)
```

To create a density plot, it compares the age who has outcome 'y' and 'n'

Calculate percentage of age

```{r}
age_percent <- marketing_data %>% group_by(age) %>% summarise(total = n(),
                                                              y_count = sum(y == 'yes'),
                                                              y_percent = (y_count/total)*100)
age_percent <- age_percent[order(age_percent$y_percent, decreasing = TRUE), ]
head(age_percent, n = 10)
```

Boxplot of age distribution to look for outliers

```{r}
ggplot(marketing_data, mapping = aes( age)) + geom_boxplot() + labs(title = "Boxplot of Age Distribution", x = "Age")+ theme_minimal() + theme( plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),)  
```

remove outliers 4100 - 4061 = 39 rows

```{r}
cleaned_data <- marketing_data %>% group_by(age) %>% filter(age < 70)
cleaned_data
```

**Duration**

```{r}
# using IQR to find outliers because the max number and the mean of duration has significant different

Q1_duration <- quantile(cleaned_data$duration, 0.25)
Q3_duration <- quantile(cleaned_data$duration, 0.75)
IQR <- Q3_duration - Q1_duration
lower_bound_duration <- Q1_duration - 1.5 * IQR
upper_bound_duration <- Q3_duration + 1.5 * IQR
print(lower_bound_duration)
print(upper_bound_duration)
print(max(cleaned_data$duration))
outliers_duration <-  cleaned_data %>% filter(duration > upper_bound_duration) %>%  count()
print(sum(outliers_duration$n)) # total outliers of duration
```

There are 295 of outliers (data transformation method is preferred)

```{r}
#(<2mins, 2-4 mins, 4-10 mins and 10 mins+)
cleaned_data <- cleaned_data %>%
  mutate(duration_category = ifelse(duration <= 120, "short", 
                             ifelse(duration > 120 & duration <= 240, "medium",
                             ifelse(duration > 240 & duration <= 600, "long", "very_long"))))

cleaned_data <- cleaned_data %>%
  mutate(duration_category = factor(duration_category, levels = c("short", "medium", "long", "very_long")))

ggplot(data = cleaned_data, aes(x = duration_category, fill = y)) + geom_bar(position = "dodge") +  labs(title = "Duration Distribution by Output",x = "Duration Category",) +  
  facet_wrap(~ y) + theme_minimal() + theme(plot.title = element_text(hjust = 0.5, size = 16, face ="bold"),)
```

Data Exploration

Repeat age distribution calculation with cleaned data

```{r}
cleaned_data <- marketing_data %>% group_by(age) %>% filter(age < 70)
ggplot(data = cleaned_data, mapping = aes(x = age, fill = y)) + geom_density(alpha = 0.6) + labs(title = "Filtered Age Distribution Vs Output") + theme_minimal() + theme(plot.title = element_text(hjust = 0.5, size = 16, face ="bold"),)
```

```{r}
cleaned_data$y <- as.factor(cleaned_data$y)

cleaned_age_percent <- cleaned_data %>% group_by(age) %>% summarise(total = n(),
         y_count = sum(y == 'yes'),  y_percent = (y_count/total))
cleaned_age_percent <- cleaned_age_percent[order(cleaned_age_percent$y_percent, decreasing = TRUE), ]
head(cleaned_age_percent, n = 10)
```

Calculate variance, standard deviation and range

```{r}
age_variance <- var(cleaned_data$age)
age_sd <- sd(cleaned_data$age)
age_range <- (cleaned_data$age)

print(age_variance)
print(age_sd)
print(age_range[2]-age_range[1])
```

Calculate again with original data

```{r}
age_variance_o <- var(marketing_data$age)
age_sd_o <- sd(marketing_data$age)
age_range_o <- (marketing_data$age)

print(age_variance_o)
print(age_sd_o)
print(age_range_o[2]-age_range_o[1])
```

Job vs age

```{r}
ggplot(cleaned_data, aes(x = age, y = job, colour = job)) + geom_point(size = 3)  +  labs(title = "Job vs age") + theme_minimal() + theme(plot.title = element_text(hjust = 0.5, size = 16, face ="bold"),)
```

Dotplot of age and education

```{r}
ggplot(data= cleaned_data, mapping = aes(x = age, fill = y) )+ geom_dotplot()+ facet_wrap(~ k, scales = "free_y") +  labs(title = "Education vs age") + theme_minimal() + theme(plot.title = element_text(hjust = 0.5, size = 16, face ="bold"),)
```

**Job**

Job Analysis after using cleaned data then it's not retired people anymore, it's student

```{r}

job_percent <- marketing_data %>% group_by(job) %>% summarise(total = n(),
                                                              y_count = sum(y == 'yes'),
                                                              y_percent = (y_count/total)*100)
job_percent <- job_percent[order(job_percent$y_percent, decreasing = TRUE), ]
head(job_percent, n = 10)
```

Using cleaned data

```{r}
job_percent <- cleaned_data %>% group_by(job) %>% summarise(total = n(),
                                                              y_count = sum(y == 'yes'),
                                                              y_percent = (y_count/total)*100)
job_percent <- job_percent[order(job_percent$y_percent, decreasing = TRUE), ]
head(job_percent, n = 10)
```

```{r}
cross_plot(cleaned_data, input = "job", target = "y")
```

**Duration**

```{r}

ggplot(data = cleaned_data, mapping = aes(x = duration, fill = y)) + labs(title = "Duration Distribution by Output") + geom_histogram(binwidth = 5) + theme_minimal() + theme(plot.title = element_text(hjust = 0.5, size = 16, face ="bold"),)
```

**Marital**

```{r}

marital_percent <- cleaned_data %>% group_by(marital) %>% summarise(total = n(),
                                                              y_count = sum(y == 'yes'),
                                                              y_percent = (y_count/total)*100)
marital_percent <- marital_percent[order(marital_percent$y_percent, decreasing = TRUE), ]
head(marital_percent, n = 10)
```

**Education (k)**

```{r}

k_percent <- marketing_data %>% group_by(k) %>% summarise(total = n(),
                                               y_count = sum(y == 'yes'),
                                               y_percent = (y_count/total)*100)
k_percent <- k_percent[order(k_percent$y_percent, decreasing = TRUE), ]
head(k_percent, n = 10)
```

```{r}
k_percent <- cleaned_data %>% group_by(k) %>% summarise(total = n(),
                                               y_count = sum(y == 'yes'),
                                               y_percent = (y_count/total)*100)
k_percent <- k_percent[order(k_percent$y_percent, decreasing = TRUE), ]
head(k_percent, n = 10)
```

**Default Credit**

```{r}
default_percent <- marketing_data %>% group_by(default) %>% summarise(total = n(),
                                               y_count = sum(y == 'yes'),
                                               y_percent = (y_count/total)*100)
default_percent <- default_percent[order(default_percent$y_percent, decreasing = TRUE), ]
head(default_percent, n = 10)
```

```{r}
default_percent <- cleaned_data %>% group_by(default) %>% summarise(total = n(),
                                               y_count = sum(y == 'yes'),
                                               y_percent = (y_count/total)*100)
default_percent <- default_percent[order(default_percent$y_percent, decreasing = TRUE), ]
head(default_percent, n = 10)
```

**housing loan**

```{r}
housing_percent <- marketing_data %>% group_by(housing) %>% summarise(total = n(),
                                               y_count = sum(y == 'yes'),
                                               y_percent = (y_count/total)*100)
housing_percent <- housing_percent[order(housing_percent$y_percent, decreasing = TRUE), ]
head(housing_percent, n = 10)
```

```{r}
housing_percent <- cleaned_data %>% group_by(housing) %>% summarise(total = n(),
                                               y_count = sum(y == 'yes'),
                                               y_percent = (y_count/total)*100)
housing_percent <- housing_percent[order(housing_percent$y_percent, decreasing = TRUE), ]
head(housing_percent, n = 10)
```

**Personal loan**

```{r}
loan_percent <- cleaned_data %>% group_by(loan) %>% summarise(total = n(),
                                               y_count = sum(y == 'yes'),
                                               y_percent = (y_count/total)*100)
loan_percent <- loan_percent[order(loan_percent$y_percent, decreasing = TRUE), ]
head(loan_percent, n = 10)
```

**Contact**

```{r}
contact_percent <- marketing_data %>% group_by(contact) %>% summarise(total = n(),
                                               y_count = sum(y == 'yes'),
                                               y_percent = (y_count/total)*100)
contact_percent <- contact_percent[order(contact_percent$y_percent, decreasing = TRUE), ]
head(contact_percent, n = 10)
```

```{r}
contact_percent <- cleaned_data %>% group_by(contact) %>% summarise(total = n(),
                                               y_count = sum(y == 'yes'),
                                               y_percent = (y_count/total)*100)
contact_percent <- contact_percent[order(contact_percent$y_percent, decreasing = TRUE), ]
head(contact_percent, n = 10)
```

**Month**

There are a lot holiday in March, Sep nad Dec

```{r}
month_percent <- marketing_data %>% group_by(month) %>% summarise(total = n(),
                                               y_count = sum(y == 'yes'),
                                               y_percent = (y_count/total)*100)
month_percent <- month_percent[order(month_percent$y_percent, decreasing = TRUE), ]
head(month_percent, n = 10)
```

```{r}
month_percent <- cleaned_data %>% group_by(month) %>% summarise(total = n(),
                                               y_count = sum(y == 'yes'),
                                               y_percent = (y_count/total)*100)
month_percent <- month_percent[order(month_percent$y_percent, decreasing = TRUE), ]
head(month_percent, n = 10)
```

**Day of Week**

```{r}
day_of_week_percent <- marketing_data %>% group_by(day_of_week) %>% summarise(total = n(),
                                               y_count = sum(y == 'yes'),
                                               y_percent = (y_count/total)*100)
day_of_week_percent <- day_of_week_percent[order(day_of_week_percent$y_percent, decreasing = TRUE), ]
head(day_of_week_percent, n = 10)
```

```{r}
day_of_week_percent$fraction = day_of_week_percent$y_percent / sum(day_of_week_percent$y_percent)
day_of_week_percent$ymax = cumsum(day_of_week_percent$fraction)
day_of_week_percent$ymin = c(0, head(day_of_week_percent$ymax, n=-1))
# Make the plot
ggplot(day_of_week_percent, aes(ymax=ymax, ymin=ymin, xmax=4, xmin=3, fill = day_of_week)) +
     geom_rect() +
     coord_polar(theta="y") + # Try to remove that to understand how the chart is built initially
     xlim(c(2, 4))  + labs(title = "Last Day of Contact") + theme_minimal()  + theme(plot.title = element_text(hjust = 0.5, size = 16, face ="bold"),)
```

**Number of perform (campaign)**

```{r}
campaign_percent <- cleaned_data %>% group_by(campaign) %>% summarise(total = n(),
                                               y_count = sum(y == 'yes'),
                                               y_percent = (y_count/total)*100)
campaign_percent <- campaign_percent[order(campaign_percent$y_percent, decreasing = TRUE), ]
head(campaign_percent, n = 10)
```

**pdays**

```{r}
pdays_percent <- marketing_data %>% group_by(pdays) %>% summarise(total = n(),
                                               y_count = sum(y == 'yes'),
                                               y_percent = (y_count/total)*100)
pdays_percent <- pdays_percent[order(pdays_percent$y_percent, decreasing = TRUE), ]
head(pdays_percent, n = 10)
```

**Number** **of contacts performed before this campaign and for this client (numeric)**

```{r}
previous_percent <- marketing_data %>% group_by(previous) %>% summarise(total = n(),
                                               y_count = sum(y == 'yes'),
                                               y_percent = (y_count/total)*100)
previous_percent <- previous_percent[order(previous_percent$y_percent, decreasing = TRUE), ]
head(previous_percent, n = 10)
```

**poutcome**

```{r}
poutcome_percent <- cleaned_data %>% group_by(poutcome) %>% summarise(total = n(),
                                               y_count = sum(y == 'yes'),
                                               y_percent = (y_count/total)*100)
poutcome_percent <- poutcome_percent[order(poutcome_percent$y_percent, decreasing = TRUE), ]
head(poutcome_percent, n = 10)
```

**job vs last day contact**

```{r}
marketing_data$day_of_week <- factor(marketing_data$day_of_week, levels = c("mon", "tue", "wed", "thu", "fri"))
ggplot(marketing_data, aes(x = day_of_week, fill = y)) + geom_bar() + facet_wrap(~job) +
  labs(title = "Last Day contacted and Job Category") + theme_minimal() + theme(plot.title = element_text(hjust = 0.5, size = 16, face ="bold"),)
```

### Correlation Analysis

**create a heatmap**

```{r}
library(corrplot)
numerical_data <- select(marketing_data, where(is.numeric))
numerical_data$output_num <- ifelse(marketing_data$y == "yes", 1,0)
corr_matrix <- cor(numerical_data, use = "complete.obs") 
corrplot(corr_matrix) 
```

```{r}
funModeling::profiling_num(cleaned_data)
```

```{r}
funModeling::plot_num(numerical_data) + theme_classic()
```

### Prepare for modelling

checking if the cleaned datset is factors

```{r}
str(cleaned_data)
```

Convert the dataset to factors

```{r}
cleaned_data[sapply(cleaned_data, is.character)] <- lapply(cleaned_data[sapply(cleaned_data, is.character)], as.factor)
cleaned_data <- cleaned_data %>% mutate(across(where(is.integer), as.numeric))
str(cleaned_data)
```

Split training

```{r}
set.seed(42)
output_idx   = createDataPartition(cleaned_data$duration, p = 0.75, list = FALSE)
output_trn = cleaned_data[output_idx, ]
output_tst = cleaned_data[-output_idx, ]
str(output_tst )
```

```{r}
dim(output_trn)
dim(output_tst)
```

### Linear Regression

```{r}
output_trn_lm = output_trn
output_tst_lm = output_tst

output_trn_lm <- output_trn_lm %>% mutate(y_num = ifelse(y == "yes", 1, 0))
output_tst_lm <- output_tst_lm %>% mutate(y_num = ifelse(y == "yes", 1, 0))

linear_model_all <- lm(y_num ~ .- y, data = output_trn_lm)

summary(linear_model_all)
```

Prediction linear regression all varaiables

```{r}
#predict
linear_pred_all <- predict(linear_model_all, newdata = output_tst)

#convert to binary class
linear_pred_class_all <- ifelse(linear_pred_all > 0.5, "yes", "no")
linear_pred_class_all <- factor(linear_pred_class_all, levels = levels(output_tst_lm$y))

# Confusion Matrix
linear_cm_all <- confusionMatrix(linear_pred_class_all, as.factor(output_tst_lm$y))

# Metrics
c(linear_cm_all$overall['Accuracy'],linear_cm_all$byClass['Pos Pred Value'],linear_cm_all$byClass['Sensitivity'])
```
```{r}
linear_accuracy <- linear_cm_all$overall['Accuracy']
linear_precision <- linear_cm_all$byClass['Pos Pred Value']  # Precision
linear_recall <- linear_cm_all$byClass['Sensitivity']
linear_F1 <-  (2 *(linear_precision * linear_recall) / (linear_precision + linear_recall))
print(linear_F1)

```

```{r}
library(pROC)

# ROC for Logistic Regression
roc_linear <- roc(output_tst$y ~ linear_pred_all, plot = TRUE, print.auc = TRUE)
plot(roc_linear, col = "blue", main = "ROC Curve for Linear Regression")
```

**AUC**

```{r}
auc_linear <- auc(roc_linear)
print(paste("AUC:", auc_linear))
```

**HTITH accuracy**

```{r}
linear_tp <- linear_cm_all$table[2, 2]
linear_fp <- linear_cm_all$table[1, 2]
linear_fn <- linear_cm_all$table[2, 1]
linear_htith <- linear_tp / (linear_tp + linear_fp + linear_fn)
linear_htith
```

```{r}
plot(linear_model_all$fitted.values, linear_model_all$residuals,
     main = "Residuals vs Fitted Values",
     xlab = "Fitted Values",
     ylab = "Residuals")
abline(h = 0, col = "red", lwd = 2)
```

### Logistic regression

Logistic Regression Model

```{r}
logistic_model_all <- glm(y ~ ., data = cleaned_data, family = binomial)
summary(logistic_model_all)
```

Predict

```{r}

# Predict on test data
logistic_pred_all <- predict(logistic_model_all, output_tst)

#convert to binary class
logistic_pred_class_all <- ifelse(logistic_pred_all > 0.5, "yes", "no")
logistic_pred_class_all <- factor(logistic_pred_class_all, levels = levels(output_tst$y))

# Confusion Matrix
logistic_cm_all <- confusionMatrix(logistic_pred_class_all, as.factor(output_tst$y))

# Accuracy, Precision, Recall
log_accuracy <- logistic_cm_all$overall['Accuracy']
log_precision <- logistic_cm_all$byClass['Pos Pred Value']  # Precision
log_recall <- logistic_cm_all$byClass['Sensitivity']       # Recall
log_F1 <- 2 * ((log_precision * log_recall) / (log_precision + log_recall))

print(logistic_cm_all)

```
```{r}
log_accuracy <- logistic_cm_all$overall['Accuracy']
log_precision <- logistic_cm_all$byClass['Pos Pred Value']  # Precision
log_recall <- logistic_cm_all$byClass['Sensitivity']     
log_F1 <-  (2 *(log_precision * log_recall) / (log_precision + log_recall))
print(log_F1)

```

ROC curve

```{r}
# Generate predicted probabilities for the "yes" class
#pred_prob <- predict(logistic_model, newdata = cleaned_data, type = #"response")

roc_logistic <- roc(output_tst$y ~ logistic_pred_all, plot = TRUE, print.auc = TRUE)

# Plot the ROC curve
plot(roc_logistic, col = "blue", main = "ROC Curve for Logistic Regression")


```

AUC

```{r}
auc_logistic <- auc(roc_logistic)
print(paste("AUC:", auc_logistic))
```

logistic htith

```{r}
log_tp <- logistic_cm_all$table[2, 2]
log_fp <- logistic_cm_all$table[1, 2]
log_fn <- logistic_cm_all$table[2, 1]
log_htith <- log_tp / (log_tp + log_fp + log_fn)
log_htith
```

k-fold cross validation (tried different method)

```{r}
# Class Predictions
# Set up cross-validation
train <- createDataPartition(cleaned_data$y, p = 0.75, list = FALSE)
data.trn <- cleaned_data[train,]
data.tst <- cleaned_data[-train,]

logistic_train <- trainControl(method = "cv", number = 50)

# Train the logistic regression model
logistic_model_train <- train(
  y ~  age + job + duration + month + euribor3m + cons.conf.idx, 
  data = cleaned_data, 
  trControl = logistic_train, 
  method = "glm", 
  family = binomial(link = "logit"), 
  metric = "Accuracy"  
)

cv_pred <- pcv_pred <- pcv_pred <- predict(logistic_model_train, data.tst)
cv_pred <- factor(cv_pred, levels = levels(data.tst$y))

# Now calculate the confusion matrix
confusionMatrix(data.tst$y, cv_pred)
```

### ***Classification tree***

```{r}
#tree_model = tree(y ~ ., data = cleaned_data)
tree_model = tree(y ~ ., data = cleaned_data, control = tree.control(nobs = nrow(cleaned_data), minsize = 10))
summary(tree_model)
```

Plot

```{r}
plot(tree_model)
text(tree_model, pretty = 0)
title(main = "Unpruned Classification Tree")
```

Prediction

```{r}

tree_pred <- predict(tree_model, newdata = output_tst, type = "class")

tree_pred <- factor(tree_pred, levels = c("no", "yes"))
output_tst$y <- factor(output_tst$y, levels = c("no", "yes"))

tree_cm <- confusionMatrix(tree_pred, output_tst$y)

print(tree_cm)

```

```{r}
c(tree_cm$overall['Accuracy'],tree_cm$byClass['Pos Pred Value'],tree_cm$byClass['Sensitivity'])
```
```{r}
tree_accuracy <- tree_cm$overall['Accuracy']
tree_precision <- tree_cm$byClass['Pos Pred Value']  # Precision
tree_recall <- tree_cm$byClass['Sensitivity']  
tree_F1 <-  (2 *(tree_precision * tree_recall) / (tree_precision + tree_recall))
print(tree_F1)
```

ROC curve

```{r}
# Train the classification tree model using rpart
tree_model_trn <- rpart(y ~ ., data = output_trn, method = "class")

tree_pred_prob <- predict(tree_model_trn, newdata = output_tst, type = "prob")

# ROC curve
#library(pROC)
roc_tree <- roc(output_tst$y, tree_pred_prob[,2], levels = c("no", "yes"))

# Plot ROC curve
plot(roc_tree, main = "ROC Curve for Classification Tree Model")
```

```{r}
auc_tree <- auc(roc_tree)
print(paste("AUC:", auc_tree))
```

tree htith

```{r}
tree_tp <- tree_cm$table[2, 2]
tree_fp <- tree_cm$table[1, 2]
tree_fn <- tree_cm$table[2, 1]
tree_htith <- tree_tp / (tree_tp + tree_fp + tree_fn)
tree_htith
```



### Clustering

**Using Hierarchical clustering**

```{r}
numerical_data <- numerical_data %>% select_if(is.numeric)
distance_matrix <- dist(scale(numerical_data), method = "euclidean")
hc <- hclust(distance_matrix, method = "complete")
plot(hc, main = "Dendrogram for Hierarchical Clustering", xlab = "", sub = "", cex = 0.6)
```

Silhouette Method

```{r}
wss <- (nrow(numerical_data) - 1) * sum(apply(numerical_data, 2, var))
wss <- sapply(1:10, function(k){
  kmeans(numerical_data, centers = k)$tot.withinss
})

plot(1:10, wss, type="b", pch = 19, frame = FALSE, 
     xlab="Number of clusters K", 
     ylab="Total within-clusters sum of squares", main = "Silhouette Method")
```

```{r}

clusters <- cutree(hc, k = 4)
numerical_data$cluster <- as.factor(clusters)
table(numerical_data$cluster)
```

```{r}
summary_by_cluster <-  numerical_data %>%
  group_by(cluster) %>%
  summarise_all(mean, na.rm = TRUE)
summary_by_cluster
```

```{r}
ggplot(numerical_data, aes(x = duration, y = nr.employed, color = cluster, main=)) + geom_point() + labs(title = "Scatter Plot of Duration vs nr.employed by Cluster") + theme_minimal()
```

**Using Ellipses method for clustering**

```{r}

featurePlot(
  x = output_trn[, c("duration", "nr.employed")],  
  y = output_trn$y,                     
  plot = "ellipse",                    
  auto.key = list(columns = 2),             
  main = "Feature Plots with Ellipses")

```
